---
title: "Project 2 Simulation Study of Cox Model and OLS Estimator"
author: "Boyi Guo"
institute: |
  | BST 765
date: "`r  Sys.Date()`"
output: 
  beamer_presentation:
    theme: "Szeged"
    colortheme: "spruce"
    toc: FALSE
    number_section: TRUE
    slide_level: 3
bibliography:
  - ref.bib
nocite: '@*'
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      # cache = TRUE,
                      message = FALSE)

library(tidyverse)
library(ggplot2)

res_30 <- readRDS("res_30.rds")
res_100 <- readRDS("res_100.rds")
```
## Objectives

* Review the simulation method for Cox proportional hazard model 
* Examine the asymptotic property of Cox proportional hazard model and least square estimator when analyzing time-to-event data (without censoring)

## Data Generating Process
### Parametric Simulation

* The survival time $y_i$ can be simulated via an exponential distribution with the rate parameter is $\beta X_i$
\begin{align*}
X_i &\sim 1 + Bernoulli(0.5), i = 1, \dots, n\\
y_i &\sim exponential (\beta X_i), i = 1, \dots, n\\
\end{align*}
* Confidence interval coverage of Cox model coefficients is not tractable in this case
* Give up

### Simulation via Inverse Hazard Function

We follow @bender2005 to simulate time-to-event data under 2-arm balanced randomized design 

* Hazard Ratio = 2
* Exponentially distributed survival time $S(t|x_i) \sim Exponential$
* No censored events
\begin{align*}
\beta & = \log(2) \quad \\
X_i & \sim Bernoulli(0.5)\\
U_i & \sim Uniform(0,1)\\
Y_i &= H_0^{-1}(-\log(U)*\exp(-\beta X_i))
\end{align*}
where the inverse function of cumulative hazard function $H_0^{-1}(x) = \lambda^{-1} x$, where $\lambda = 0.5$

### Simulation Parameters
* Power calculation^[http://powerandsamplesize.com/Calculators/Test-Time-To-Event-Data/Cox-PH-2-Sided-Equality]

  * Assume $\alpha = 0.05$
  * power ( 1 - $\beta$) = 0.80 $\Rightarrow$ $N_{total} = 66$, i.e. 33 subjects per group
  * $N_{total}  = 30 \Rightarrow$  power ( 1 - $\beta$) $\approx$ 0.48 
  * $N_{total}  = 100 \Rightarrow$  power ( 1 - $\beta$) $\approx$ 0.93
  
* $n_{it} = 1000$ replications

## Result

## Cox Proportional Hazard Models
```{r}
ggplot() +
  # geom_histogram(aes(x = exp(cox_coef), color = "N=30", fill = "N=30"), data = res_30, alpha = 0.5) +
  # geom_histogram(aes(x = exp(cox_coef), color = "N=100", fill = "N=100"), data = res_100, alpha = 0.5) +
  geom_density(aes(x = cox_coef, color = "N=30", fill = "N=30"), data = res_30, alpha = 0.5) +
  geom_density(aes(x = cox_coef, color = "N=100", fill = "N=100"), data = res_100, alpha = 0.5) +
  geom_vline(xintercept = log(2), linetype = 2) +
  # geom_vline(xintercept = mean(res_30$cox_coef)) +
  # geom_vline(xintercept = mean(res_100$cox_coef))
  ylab("Density")+
  xlab("Log Hazard Ratio") +
  ggtitle("Sampling Distribution of Log Hazard Ratio") +
  theme_classic()
```

* Reject rate:
  * 

## Least Square Estimator
```{r}
ggplot() +
  # geom_histogram(aes(x = exp(cox_coef), color = "N=30", fill = "N=30"), data = res_30, alpha = 0.5) +
  # geom_histogram(aes(x = exp(cox_coef), color = "N=100", fill = "N=100"), data = res_100, alpha = 0.5) +
  geom_density(aes(x = ols_coef, color = "N=30", fill = "N=30"), data = res_30, alpha = 0.5) +
  geom_density(aes(x = ols_coef, color = "N=100", fill = "N=100"), data = res_100, alpha = 0.5) +
  ylab("Density")+
  xlab("OLS Estimate") +
  ggtitle("Sampling Distribution of Ordinary Least Square Estimator") +
  theme_classic()
```


### Coverage and Rejection Rate
* Cox model rejection rate and confidence interval is as expected. 
  * The coverage is slightly smaller than the anticipated 0.95 possibly due to simulation random error or Wald test is too liberal
  * Suggest to run with larger number of replications
  
* Supprisingly, OLS model have similar rejection rate as Cox model. 

```{r}
rbind(
res_30 %>% 
  summarize(cox_rej = sum(cox_p < 0.05)/n(),
            ols_rej = sum(ols_p < 0.05)/n(),
            cox_cov = sum(cox_CI.97.5.. > log(2) & cox_CI.2.5.. < log(2))/n()) %>% 
  mutate(N = 30),
res_100 %>% 
  summarize(cox_rej = sum(cox_p < 0.05)/n(),
            ols_rej = sum(ols_p < 0.05)/n(),
            cox_cov = sum(cox_CI.97.5.. > log(2) & cox_CI.2.5.. < log(2))/n()) %>% 
  mutate(N = 100)
) %>% 
  select(N, cox_cov, cox_rej, ols_rej) %>% 
  knitr::kable(col.names = c("Sample Size", "Cox Coverage", "Cox Rejection", "OLS Rejection"))
```



## Other Survial Data Simulations

* @crowther2012 introduced time-to-event simulation when the baseline hazard function is more complicated or competing risk presents 

* @austin2012 introduced survival data simulation when time-varying covariates present

## Reference