---
title: "Project 1 Evaluate RNG"
author: "Boyi Guo"
institute: |
  | BST 765
date: "Feb 16, 2021"
output: 
  beamer_presentation:
    theme: "Szeged"
    colortheme: "spruce"
    toc: FALSE
    number_section: false
    slide_level: 2
bibliography:
  - ref.bib
nocite: '@*'
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      cache = TRUE,
                      message = FALSE)

if(!require("e1071")) install.packages("e1071")
library(dplyr)
library(ggplot2)
library(knitr)
library(gridExtra)
library(grid)

# Transformation Constants
c0 <- 0.108304
c1 <- 1.104252
c2 <- -0.123347
c3 <- -0.045284
c4 <- 0.005014
c5 <- 0.001285

# Simulation Parameters
n_sample <- 1e5
n_it <- 1000

# Ancillary functions
beta_mean <- function(alpha, beta) alpha/(alpha+beta)
beta_var <- function(alpha, beta) alpha*beta/((alpha+beta)^2*(alpha+beta+1))
beta_sd <- function(alpha, beta) sqrt(beta_var(alpha, beta))
beta_skew <- function(alpha, beta)
  (2*(beta-alpha)*sqrt(alpha+beta+1))/((alpha+beta+2)*sqrt(alpha*beta))
beta_kurtosis <- function(alpha, beta)
  6*((alpha-beta)^2*(alpha+beta+1)-alpha*beta*(alpha+beta+2))/(alpha*beta*(alpha+beta+2)*(alpha+beta+3))
```

## Data Generating Process

We follow @headrick2002 to simulate a Beta distribution ($\alpha = 4, \beta = 2$) from a normal distribution with mean 0 and standard deviation 1. 

\begin{equation}
  \label{eq:transf}
  Y_Z = c_0 + c_1 X + c_2X^2 + c_3X^3 + c4X^4 + c_5X^5,
\end{equation}
\begin{equation}
  \label{eq:scale}
  Y = Y_Z*\sigma_{\beta_{4,2}} + \mu_{\beta_{4,2}},
\end{equation}
where $X$ follows a standard normal distribution, N(0,1), $c_0 = `r c0`, c_1 = `r c1`, c_2 = `r c2`, c_3 = `r c3`, c_4 = `r c4`, c_5 = `r c5`$. 

## Simulation
* Simulate $X_i, i=1,\dots,`r (n_sample <- 1e5) %>% format(scientific=FALSE, big.mark = ",")` \sim N(0,1)$ 
* $Y_i = f_{Hendrix}(X_i)$ described in Equation 1, 2
* Repeat `r n_it` iterations 

The simulation is conducted on a 64-bit Windows 10 Platform machine with Inter i5 processor and 8 GB RAM. The simulation is implemented in `r R.version$version.string`.


```{r echo = F}
# Note: the constants c have been loaded in the computation environment
hendrick_beta_4_2 <- function(x) {
  ret <- c0 + c1*x + c2*(x^2) + c3*(x^3) + c4*(x^4) + c5*(x^5)
  # return to the original scale and position
  ret*beta_sd(4,2) + beta_mean(4,2)
}

# wrapper function for each iteration in the simulation
sim_iteration <- function(
  it,
  n_sample,  # Sample Size
  func       # Transformation Function
){
  # Simulate X
  X <- rnorm(n_sample, mean = 0, sd = 1)
  Y <- func(X)
  
  # return
  data.frame(
    it = it,
    mean = mean(Y),
    var = var(Y),
    skew = e1071::skewness(Y),
    kurt = e1071::kurtosis(Y, type = 1)
  )
}

# Simulation Body
set.seed(1)

sim_res <- purrr::map_dfr(1:n_it, .f = sim_iteration,
                          n_sample = n_sample, func = hendrick_beta_4_2)
```

## Result
For Beta distribution with parameters ($\alpha, \beta$), the moments can be calculated following the equations below:
\begin{align*}
  \mu & = \frac{\alpha}{\alpha+\beta} \\
  \sigma^2 &= \frac{\alpha*\beta}{(\alpha+\beta)^2*(\alpha+\beta+1)} \\
  \text{skewness} & = 
  \frac{2*(\beta-\alpha)*\sqrt(\alpha+\beta+1)}{(\alpha+\beta+2)*\sqrt(\alpha*\beta)}\\
  \text{excess kurtosis}&  = \frac{6*((\alpha-\beta)^2(\alpha+\beta+1)-\alpha*\beta*(\alpha+\beta+2))}{\alpha*\beta*(\alpha+\beta+2)*(\alpha+\beta+3)}.
\end{align*}

## Result (Cont.)
* The averaged observed moments match with the epxected moments closesly, up to 4 digits. 

```{r echo = FALSE}
expct <- c(Mean = beta_mean(4,2),
           Variance = beta_var(4,2),
           Skewness = beta_skew(4,2),
           "Excess Kurtosis" = beta_kurtosis(4,2))

sim_sum = sim_res %>% 
  select(-it) %>% 
  summarize_all(mean)

rbind("Expected" = expct, 
      "Observed" = sim_sum) %>% 
  kable(
    col.names = c("Mean", "Variance", "Skewness", "Excess Kurtosis"),
    align = "lcccc",
    caption = "Expected and observed moments of Beta(4,2)",
    ref = "tab:comparison"
  )

```
## Result (Cont.)
* Central Limit effect exhibits, especially in variance and skewness.

```{r echo = FALSE, fig.height= 5}
plot_mean <- ggplot(sim_res) +
  geom_histogram(aes(x = mean, y=..density..),
                 color="black", fill="lightblue") +
  geom_density(aes(x = mean), size = 1.5) +
  xlab("Mean") +
  ylab("Density") +
  theme_classic()

plot_var <- ggplot(sim_res) +
  geom_histogram(aes(x = var, y=..density..),
                 color="black", fill="lightblue") +
  geom_density(aes(x = var), size = 1.5) +
  xlab("Variance") +
  ylab("Density") +
  theme_classic()
  
plot_skew <- ggplot(sim_res) +
  geom_histogram(aes(x = skew, y=..density..),
                 color="black", fill="lightblue") +
  geom_density(aes(x = skew), size = 1.5) +
  xlab("Skewness") +
  ylab("Density") +
  theme_classic()

plot_kurt <- ggplot(sim_res) +
  geom_histogram(aes(x = kurt, y=..density..),
                 color="black", fill="lightblue") +
  geom_density(aes(x = kurt), size = 1.5)  +
  xlab("Kurtosis") +
  ylab("Density") +
  theme_classic()

grid.arrange(plot_mean, plot_var, plot_skew, plot_kurt,
             ncol=2,
               top = textGrob("Sample Distribution of Moments"))

```

## Goodness of fit

 * there are data points out of the range of a Beta distribution. 

```{r  echo = FALSE, fig.height= 5}
set.seed(1)
set.seed(sample(1:100,1))
X <- rnorm(n_sample, mean = 0, sd = 1)
Y <- hendrick_beta_4_2(X)
  
ggplot() +
  geom_histogram(aes(x = Y, y=..density..),
                 color="black", fill="lightblue") +
  geom_density(aes(x = Y, color = "Empirical"), size = 1.5, alpha = 0.25)+
  geom_function(aes(color = "Expected"), size = 1.5, alpha = 0.25,
                fun = dbeta, args = list(shape1 = 4, shape2 = 2)) +
  theme_classic() +
  ylab("Density") +
  xlab("Value")


```


\newpage
## Reference